UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-07-04 ms/main.tex
fix some images, add more stuff on FEV1, add citation on AFR/EUR FEV1 study

Forced expiratory volume in 1 second (FEV1) (figures~\ref{fig:supp_ukbb_fev_f}, \ref{fig:supp_ukbb_fev_m}) also shows strong correlations with certain populations --- South Asian, African, and Caribbean --- having considerably lower measurements on average (see figures~\ref{fig:supp_box_fev_f} and \ref{fig:supp_box_fev_m} for boxplots and p-values). Notably, there appears to be a juncture in the admixture population, highlighted in figure~\ref{fig:supp_comparison_fev_afr}, where the distribution of FEV1 changes. This roughly corresponds to the transition from Black African/Caribbean individuals to those who identified having mixed backgrounds. Boxplots and statistical testing suggest that relative to White British populations, FEV1 values are significantly lower for Black African and Black Caribbean populations, but not for White and Black Caribbean and White and Black African populations. Figure~\ref{fig:supp_comparison_fev_chi_eur} suggests a difference in FEV1 between females who self-identified as Chinese and a nearby cluster enriched in individuals born in Japan. An unpaired two-sample t-test shows those born in Japan have a higher mean FEV1 by 0.236 (p-value $4.902\times 10^{-11}$); adjusting for age, age$^2$, height, and sex shows a difference of 0.213 (p-value $5.4\times 10^{-13}$) among females and 0.317 (p-value $5.1\times 10^{-4}$) among males (distributions presented in figure~\ref{fig:supp_fev_ridgeplots}).

>>>>

Forced expiratory volume in 1 second (FEV1) (figures~\ref{fig:supp_ukbb_fev_f}, \ref{fig:supp_ukbb_fev_m}) also shows strong correlations with certain populations --- South Asian, African, and Caribbean --- having considerably lower measurements on average (see figures~\ref{fig:supp_box_fev_f} and \ref{fig:supp_box_fev_m} for boxplots and p-values). Notably, there appears to be a juncture in the admixture population, highlighted in figure~\ref{fig:supp_comparison_fev_afr}, where the distribution of FEV1 changes. This roughly corresponds to the transition from Black African/Caribbean individuals to those who identified having mixed backgrounds. Boxplots and statistical testing suggest that relative to White British populations, FEV1 values are significantly lower for Black African and Black Caribbean populations, but not for White and Black Caribbean and White and Black African populations. Figure~\ref{fig:supp_comparison_fev_chi_eur} suggests a difference in FEV1 between those who self-identified as Chinese and a nearby cluster enriched in individuals born in Japan. Though studies in several populations have shown an inverse relationship between African ancestry and FEV1 levels compared to European populations\cite{ortega2015effect}, we are not aware of such results for specific Asian ancestries. We selected the individuals whose UMAP coordinates were near the Chinese cluster and kept those born in Japan, Malaysia, and the Philippines (the largest populations in the UKBB by country of birth) as well as the self-identified Chinese population (mutually exclusive groups, shown in figure~\ref{fig:supp_fev_test_pops}). After adjusting for age, age$^2$, height, and sex, an unpaired two-sample t-test shows those born in Japan have a higher mean FEV1 than Chinese individuals by 0.224 (p-value $2.787\times 10^{-15}$). By sex, there is a difference of 0.213 (p-value $5.4\times 10^{-13}$) among females and 0.317 (p-value $5.1\times 10^{-4}$) among males, though there are considerably fewer males in the sample (distributions presented in figure~\ref{fig:supp_fev_ridgeplots}). For comparison, the adjusted difference between self-identified African and British individuals is 0.762 (p-value $2.2\times 10^{-16}$).
